### 2021-11-29 plos_latex_template.tex
Minor changes to Intro & Pillars ~ @jananiravi

% Rocío: Main message of the paragraph: Conferences are great, but for some
Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
The larger the conference, the larger the opportunities to network and learn from the cohort.
However, opportunities for participating in conferences are not equal for all. 
Just like in academia and tech in general, conferences can become exclusive spaces for privileged groups, i.e. with unearned advantages given by society to some but not all (e.g. white, male, from a rich country, English-native speaker, with no physical disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
% However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.

This article proposes 10 rules to host inclusive conferences.
They are organized in three sections: the pillars (embracing diversity, creating a welcoming environment, and having a diverse organizing team), design rules (counteracting bias in the conference program, having a strong online component, achieving accessibility for people with disabilities, embracing language diversity, having inclusive communication strategies, and allocating financial resources to support inclusive practices), and a process rule: make the conference part of a long-term process for inclusion. 
These practices stem from the experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible.

The list of authors is composed of the global coordinators of the conference, members of the diversity and inclusion team, and people who contributed to other areas of the conference. 
Therefore, the rules are a result of our lessons learned from diverse perspectives and expertise (i.e. our different backgrounds, profiles, and roles in the conference). 
Even if this set of rules is derived from our experience with a global scale event, we believe these can be easily transferred to events with regional or local scopes.
They are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
The rules can also be helpful to local/virtual organizers who intend to put inclusion at the core of the planning phase.

@@

\section*{The pillars}

@@

% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
Human diversity should be celebrated and respected in every way. Nonetheless, we live in a world with implicit hierarchies along these axes. Some statuses (\textit{e.g.}, cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems--including conferences--are consciously and unconsciously built. \as{People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia. No form of oppression operates in isolation from the others, which means some groups are particularly vulnerable and systematically left behind.}
While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}. 
Recognizing this systematic exclusion, particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against.
These are the groups we need to make more effort to include.
Investing more effort towards the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
% vision
% back to the indicators? 
% realistic but bold (back?)
% be aware and promote awareness
%¡LAS BARRERAS!
%Express the diversity and inclusion vision for your conference by writing a diversity statement.

% Rocío: more effort towards including the more excluded
This may translate into different actions depending on your field, region, or community, e.g. a gender distribution of your speakers that is representative of the general population, presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.

@@

\section*{Design Rules}

@@

\section*{The Process}

>>>>

% Rocío: Main message of the paragraph: Conferences are great, but for some
Conferences, from the Latin \textit{conferre} `to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
The larger the conference, the larger the opportunities to network and learn from the cohort.
However, opportunities for participating in conferences are not equal for all. 
Although unintentionally, many academic and tech conferences have become exclusive spaces for privileged groups, \textit{i.e.}, with unearned advantages given by society to some but not all (\textit{e.g.}, white, male, from a rich country, English-native speaker, with no physical disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
% However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.

This article proposes 10 simple rules to host inclusive conferences.
They are organized in three sections: \textbf{the pillars} (embracing diversity, creating a welcoming environment, and having a diverse organizing team), \textbf{design rules} (counteracting bias in the conference program, embracing a strong online component, achieving accessibility for people with disabilities, embracing language diversity, enforcing inclusive communication strategies, and allocating financial resources to support inclusive practices), and a \textbf{process rule} to make the conference part of a long-term commitment to inclusion. 
These practices stem from the experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible.

The authors of this article include the global coordinators of the conference, members of the diversity and inclusion team, and people who contributed to other critical areas of the conference. 
Therefore, the rules result from our lessons learned from diverse perspectives and expertise (\textit{i.e.}, our different backgrounds, profiles, and roles in the conference). 
Even though these rules were designed based on our experience with a global event, they can be easily transferred or scaled to events with regional or local scopes.
These rules will greatly benefit people who are part of meeting committees that oversee the site/location selection process, or that coordinates with the local organizers of conferences.
The rules will also help local/virtual organizers who intend to put inclusion at the core of their conference right from the conception and planning phases. We have designed them for wide adoption to enable and ensure that the scientific and technical conferences post-pandemic are as accessible and inclusive as they can be.

@@

\section*{I. The pillars}

@@

% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
Human diversity should be celebrated and respected in every way. Nonetheless, we live in a world with implicit hierarchies along these axes. Some statuses (\textit{e.g.}, cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems--including conferences--are consciously and unconsciously built. \as{People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia. No form of oppression operates in isolation from the others, which means some groups are particularly vulnerable and systematically left behind.}
While no isolated initiative can drastically alter conference outcomes, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}. 
Recognizing this systematic exclusion, particularly in our field and in our scientific or professional community, will help identify the most marginalized and discriminated subgroups.
These are the groups that warrant deliberate efforts that ensure fair inclusion.
Investing a substantial effort towards the most excluded groups does not mean neglecting the other communities, but will instead ensure equity, guide the vision of diversity for your conference, and help design strategies to achieve it.
% vision
% back to the indicators? 
% realistic but bold (back?)
% be aware and promote awareness
%¡LAS BARRERAS!
%Express the diversity and inclusion vision for your conference by writing a diversity statement.

% Rocío: more effort towards including the more excluded
This may translate into different actions depending on your field, region, or community, \textit{e.g.}, a gender distribution of your speakers that is representative of the general population, presence of racialized people — especially Black people — among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.

@@

\section*{II. Design Rules}

@@

\section*{III. The Process}
